Testing Quasar Unification with Clumpy Wind Models

### 2015-12-09 paper.tex
still working
reordered paper

%% Lines and ions
\def\civ{C~\textsc{iv}}
\def\nv{N~\textsc{v}}
\def\hei{He~\textsc{i}}
\def\heii{He~\textsc{ii}}
\def\mg{Mg~\textsc{ii}}
\def\al{Al~\textsc{iii}}
\def\heii{He~\textsc{ii}}
\def\ovi{O~\textsc{vi}}
\def\la{Ly~$\alpha$}
\def\ha{H~$\alpha$}
\def\hb{H~$\beta$}

@@

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind. 
According to this paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows  in radio-loud (RL) sources \citep{zhou2006}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.

Due to their ubiquitous nature,
disc winds offer a natural explanation for the
diverse phenemonology of luminous AGN and quasars \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this {\em geoemtric unification} framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

\subsection{Ionization treatment}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
This allows the calculation of a photoionization rate estimator. Ion abundances are
then calculated by solving the rate equations between ions. We include collisional ionization
and photoionization balanced with radiative, 
dielectronic and collisional (three-body) recombination.
As in M15, we use a dilute Boltzmann approximation to calculate 
the population of levels for simple-ions. We stress that this approximation 
is not required for ions treated as macro-atoms. 
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
% The ionization rate out of ion $j$ can then be written as 
% \begin{equation}
% R_{j,j+1}(J)= 
% \displaystyle{
% n_j \left(C_{j} n_e + 
% \sum_{band~i=0}^{n}~
% {\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
% \right),}
% \end{equation}
% where $\sigma_j$ is the photoionization cross-section, $n_e$
% is the electron density and $C_{j}$
% represents the collisional ionization coefficient.
% The recombination rate {\em into} ion $j$ is given by 
% \begin{equation}
% R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
% \end{equation}
% where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
% The subscripts denote radiative and 
% collisional (three-body) recombination, respectively. 
% $T_e$ is the electron temperature.
% %% We neglect recombination to and from excited states in the simple-ion calculation. 
% For simple-ions, we use a dilute Boltzmann approximation to calculate 
% the population of level $k$ in ionic stage $j$,
% \begin{equation}
% \frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
% \end{equation}
% Here $z_j$ is the partition function of ionic stage $j$,
% $T_R$ is the effective radiation temperature,
% $E_k$ is the energy difference between level $k$ and the ground state,
% and $g_k$ is the statistical weight of level $k$. 
% We stress that this approximation is not required for ions
% treated as macro-atoms. 
% {\bf KSL: 
% It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.
% JM: We've shortened the section on simple-ions as agreed..}

@@

\subsection{The Simulation Grid: Arriving at a next-generation model}

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la?
\item Do the spectra  produce LoBAL features at a  small subset of BAL angles?
\item Does the model spectra resemble quasar composite spectra?
\end{itemize}
In the next section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
We then discuss the sensitivity to key parameters in section 5.
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\section{Results and Discussion}

Here we describe the results from our next-generation model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. We also examine trends with inclination in the synthetic spectra, 
both in terms of the range of ionization states of the absorption lines and equivalent widths 
of the emission lines.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure*} %fullpage

@@

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
density squared. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities 
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
represent the Sobolev escape probability, upper level number density and 
Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles in our model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
XXXXX (REF), respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled using a continuum fit
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*} %fullpage

@@

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999a,arav1999b}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{surlan2012,hamann2008}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.  
% {\bf KSL: it sounds like what you are suggesting is a low density inner flow, our domain 2 stuff actually permits ... or indeed one might be able to produce this within SV is you adopted a different mass loss with radius assumption. JM: I am being deliberately 
% vague about what I am suggesting. I am saying there are three physical
% possibilities: either wind itself causes this scattering contribution, which doesn't happen in our model, or you have partial covering, or you have some additional scatterer.}

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/ew.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:ew}
\end{figure*} %fullpage

@@

\subsubsection{The Angular Distribution of Line And Continuum Emission}
\label{{sec:angular}}

In order to quantitatively assess how emission lines change with 
inclination when blue-shifted absorption 
may affect the line profile, we define the `red wing equivalent width' ($W_{\lambda,RW}$) as
\begin{equation}
W_{\lambda,RW} = \int_{\lambda_0}^{\lambda^\prime} \left( 1 - \frac{F_\lambda}{F_0} \right) d\lambda
\label{rwew}
\end{equation}
where $F_0$ is the continuum flux and the integral is calculated from $\lambda_0$, line centre,
to a wavelength $\lambda^\prime$ where the flux has returned to the continuum level.
This quantity is shown as a function of inclination in figure 5 for the \civ\ and \mg\ UV lines.
We also plot show the $W_{\lambda,RW}$ expected from isotropic line emission and a foreshortened 
and limb darkened disc  as well as $1/2$ equivalent widths from \cite{dipompeo2012b}.

@@

It is also possible that the equatorial paradigm invoked from early polarisation
studies \citep{goodrich1995, cohen1995,brotherton2006} 
is an over-simplification, or is incorrect. 
High brightness temperatures in some RL BALQSOs imply polar outflows \citep{zhou2006} 
and \cite{bruni2012} find that
RL BALQSOs possess similar radio spectral indices to normal RL quasars, 
suggestive of comparable inclinations. 
In addition, \cite{marin2013} find a bending angle of $\sim45^\circ$ is required to 
explain the polarisation dichotomy of type 1 and 2 AGN using an Elvis-type wind 
model (Elvis 2000). It is therefore possible 
that type 1 quasars and BALQSOs are generally viewed from a fairly narrow range of angles 
($\sim0$-$45^\circ$), or that {\em both} evolutionary and geometric explanations are required.
We suggest that future modelling should include predictions of polarisation signatures
from a detailed radiative transfer simulation, allowing direct comparison with spectropolarimetry
of BALQSOs.

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/lx_a05_pre.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

The main motivation for including a treatment of clumping was
to avoid over-ionization of the wind in the presence of strong X-rays. 
Having verified that strong BALs appear in the synthetic spectra,
it is also important to assess whether the X-ray properties of this
next-generation model agree well with quasar and BALQSO samples for the relevant
inclinations.

Figure~\ref{fig:xray} shows the emergent
monochromatic luminosity ($L_\nu$) at 2~keV and 
plotted against $L_\nu$ at $2500$\AA\ for a number of different viewing angles in our model.
The monochromatic luminosities are calculated from the synthetic spectra and thus include
the effects of wind reprocessing and attenuation. In addition to model outputs,
we also show the BALQSO sample of Saez et al. (2012) and luminous AGN and quasar
samples from Steffen et al. (2006). The best fit relation from Steffen et al. (2006) 
is also shown. For low inclination, `quasar-like' viewing angles,
we now find excellent agreement with AGN samples. The gradient from $20^\circ$ to
$60^\circ$ in our models is caused by a combination of disc foreshortening/limb-darkening 
(resulting in a lower $L_{2500}$ for higher inclinations) and the fact that the disk 
is opaque, and thus the X-ray source subtends a smaller solid angle at high inclinations
(resulting in a lower $L_{2keV}$ for higher inclinations).

Our models slightly over-predict the emergent X-ray luminosity at BAL angles, 
although we are limited by poor sample sizes. 
% It is possible that this is due to our wind being overly optically thick to 
% electron scattering at angles which look through the wind. 
If BALQSOs were {\em intrinsically} 
X-ray weak \citep[as suggested by, e.g.][]{morabito2013},
our isotropic assumption for the X-ray source would be incorrect. 
A polar-biased X-ray source would result in a lower clumping factor being
required in our model. Our specific wind
prescription will also affect the opacities, densities and resultant
ionization structure, which can change the absorption characteristics and resultant
luminosities.
Nevertheless, our input X-ray spectrum
now reproduces the X-ray properties of a luminous quasar as an output,
and at least some BAL angles match the observations.
This satisifies the first-order requirement for the X-ray properties of 
a unified quasar model.

>>>>

%% Lines and ions
\def\civ{C~\textsc{iv}}
\def\nv{N~\textsc{v}}
\def\hei{He~\textsc{i}}
\def\heii{He~\textsc{ii}}
\def\heiiline{He~\textsc{ii}~$4686$\AA}
\def\mg{Mg~\textsc{ii}}
\def\al{Al~\textsc{iii}}
\def\heii{He~\textsc{ii}}
\def\ovi{O~\textsc{vi}}
\def\la{Ly~$\alpha$}
\def\ha{H~$\alpha$}
\def\hb{H~$\beta$}

@@

Perhaps the clearest evidence of outflows in AGN is  
the blueshifted ($\sim 0.1c$) broad absorption lines (BALs) in the 
ultraviolet seen in approximately $20\%$ of quasars
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
The simplest explanation for the incidence of 
BAL quasars (BALQSOs) is in terms of an accretion disc wind. 
According to this paradigm, a biconical wind rises from 
the accretion disc and the BALQSO fraction is associated with
the covering factor of the outflow. 
Polarisation studies expect the wind to be roughly equatorial
\citep{goodrich1995, cohen1995}, although there is also evidence
for polar BAL outflows  in radio-loud (RL) sources \citep{zhou2006,ghoshpunsly2007}.
% although this geometry is contradicted
% by more recent studies of radio spectral indices \citep{dipompeo2012}.

Due to their ubiquitous nature,
disc winds offer a natural explanation for the
diverse phenemonology of luminous AGN and quasars \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this {\em geometric unification} framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

\subsection{Ionization and Excitation Treatment}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
This allows the calculation of a photoionization rate estimator. Ion abundances are
then calculated by solving the rate equations between ions. We include collisional ionization and photoionization balanced with radiative, 
dielectronic and collisional (three-body) recombination.
As in M15, we use a dilute Boltzmann approximation to calculate 
the populations of levels for simple-ions. We stress that this approximation 
is not required for macro-atom levels. 
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
% The ionization rate out of ion $j$ can then be written as 
% \begin{equation}
% R_{j,j+1}(J)= 
% \displaystyle{
% n_j \left(C_{j} n_e + 
% \sum_{band~i=0}^{n}~
% {\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
% \right),}
% \end{equation}
% where $\sigma_j$ is the photoionization cross-section, $n_e$
% is the electron density and $C_{j}$
% represents the collisional ionization coefficient.
% The recombination rate {\em into} ion $j$ is given by 
% \begin{equation}
% R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
% \end{equation}
% where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
% The subscripts denote radiative and 
% collisional (three-body) recombination, respectively. 
% $T_e$ is the electron temperature.
% %% We neglect recombination to and from excited states in the simple-ion calculation. 
% For simple-ions, we use a dilute Boltzmann approximation to calculate 
% the population of level $k$ in ionic stage $j$,
% \begin{equation}
% \frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
% \end{equation}
% Here $z_j$ is the partition function of ionic stage $j$,
% $T_R$ is the effective radiation temperature,
% $E_k$ is the energy difference between level $k$ and the ground state,
% and $g_k$ is the statistical weight of level $k$. 
% We stress that this approximation is not required for ions
% treated as macro-atoms. 
% {\bf KSL: 
% It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.
% JM: We've shortened the section on simple-ions as agreed..}

@@

\subsection{The Simulation Grid}
% Arriving at a next-generation model}

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}).

We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)?
\item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ \citep[e.g. ][]{shen2011}?
\item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la 
\citep[e.g. ][]{dipompeo2012}?
\item Do the spectra produce \mg\ and \al\ (LoBAL) absorption 
features with $BI > 0$ at a small subset of BAL angles?
\item Does the model spectra resemble quasar composite spectra?
\end{itemize}
Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

In the next section, we present one of the most promising models,
which we refer to as the fiducial model, and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
We then discuss the sensitivity to key parameters in section 5.
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\section{Results From A Fiducial Model}

Here we describe the results from an fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities, $L$, represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure} %fullpage

@@

The total line luminosity also increases dramatically compared to H13. 
This is due to heating of the plasma by the increased the increased X-ray 
luminosity heating the plasma, while clumping prevents over-ionization.
This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the incident X-rays.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities, $L$,
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
% This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
% represent the Sobolev escape probability, upper level number density and 
% Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles in our model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
XXXXX (REF), respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled using a continuum fit
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure} %fullpage

@@

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999b,arav1999}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{hamann2008,surlan2012}.
Second, our wind does not have a significant scattering 
contribution along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc.  
% {\bf KSL: it sounds like what you are suggesting is a low density inner flow, our domain 2 stuff actually permits ... or indeed one might be able to produce this within SV is you adopted a different mass loss with radius assumption. JM: I am being deliberately 
% vague about what I am suggesting. I am saying there are three physical
% possibilities: either wind itself causes this scattering contribution, which doesn't happen in our model, or you have partial covering, or you have some additional scatterer.}

@@

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/lx_a05_pre.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure} %fullpage

The main motivation for including a treatment of clumping was
to avoid over-ionization of the wind in the presence of strong X-rays. 
Having verified that strong BALs appear in the synthetic spectra,
it is also important to assess whether the X-ray properties of this
fiducial model agree well with quasar and BALQSO samples for the relevant
inclinations.

Figure~\ref{fig:xray} shows the emergent
monochromatic luminosity ($L_\nu$) at 2~keV and 
plotted against $L_\nu$ at $2500$\AA\ for a number of different viewing angles in our model.
The monochromatic luminosities are calculated from the synthetic spectra and thus include
the effects of wind reprocessing and attenuation. In addition to model outputs,
we also show the BALQSO sample of Saez et al. (2012) and luminous AGN and quasar
samples from Steffen et al. (2006). The best fit relation from Steffen et al. (2006) 
is also shown. For low inclination, `quasar-like' viewing angles,
we now find excellent agreement with AGN samples. The slight gradient from $20^\circ$ to
$60^\circ$ in our models is caused by a combination of disc foreshortening/limb-darkening 
(resulting in a lower $L_{2500}$ for higher inclinations) and the fact that the disk 
is opaque, and thus the X-ray source subtends a smaller solid angle at high inclinations
(resulting in a lower $L_{2keV}$ for higher inclinations).

% JM XXX -- only include this if it's true!
% Our models slightly over-predict the emergent X-ray luminosity at BAL angles, 
% although we are limited by poor sample sizes. 
% % It is possible that this is due to our wind being overly optically thick to 
% % electron scattering at angles which look through the wind. 
% If BALQSOs were {\em intrinsically} 
% X-ray weak \citep[as suggested by, e.g.][]{morabito2013},
% our isotropic assumption for the X-ray source would be incorrect. 
% A polar-biased X-ray source would result in a lower clumping factor being
% required in our model. Our specific wind
% prescription will also affect the opacities, densities and resultant
% ionization structure, which can change the absorption 
% characteristics and resultant luminosities.
% Nevertheless, our input X-ray spectrum
% now reproduces the X-ray properties of a luminous quasar as an output,
% and at least some BAL angles match the observations.
% This satisifies the first-order requirement for the X-ray properties of 
% a unified quasar model.

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/c4_angles.png}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for viewing angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure} %fullpage

At certain sightlines, the synthetic spectra exhibit blue-shifted BALs in \al\ and \mg --
the absorption lines seen in LoBALQSOs, and we even see absorption in Fe~\textsc{ii}
at the highest inclinations. Line profiles in velocity space 
for \civ, \al\ and \mg, are shown in figure~\ref{fig:lobal} for a range
of BALQSO viewing angles. We find that ionization stratification
of the wind causes lower ionization material to have a smaller covering factor, 
as demonstrated by figures~\ref{fig:wind} and \ref{fig:lobal}.
This confirms the behaviour expected from a unification model such as Elvis (2000). 
LoBALs are only present at viewing angles close to edge-on ($i>75^\circ$),
as predicted by polarisation results \citep{brotherton1997}.
As observed in a BALQSO sample by \cite{filizak2014}, we find that
BAL troughs are wider and deeper when low ionization absorption features are present,
and high ionization lines have higher blue-edge velocities than the 
low ionization species.
There is also a correlation between the strength of LoBAL features
and the amount of continuum attenuation at that sightline, particularly
blueward of the Lyman edge as the low ionization base 
intersects the line-of-sight. 
A model such as this therefore predicts that LoBALQSOs and FeLoBALQSOs 
have stronger Lyman edge absorption and 
are more Compton-thick than HiBALQSOs and Type 1 quasars.
An edge-on scenario also offers a potential explanation for the rarity of LoBAL and
FeLoBAL quasars, due to a foreshortened and attenuated continuum, 
although, as noted in section~\ref{angular}, BAL fraction 
inferences are fraught with complex selection effects.

% DISCUSSION???

 \section{Discussion: Parameter Sensitivity and Inclination Dependence}

 We have demonstrated that our fiducial model successfully reproduces much of
 the phenomenology of luminous AGN and quasars, with a few exceptions.
 Here, we discuss these shortcomings, by discussing parameter sensitivity
 and the dependence of line EWs on inclination.

\subsection{The Angular Distribution of Line And Continuum Emission}
\label{sec:angular}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/ew.png}
\caption
{
$W_{\lambda,RW}$ as a function of inclination in our example model (red line).
We also the mean $1/2$ equivalent width from DiPompeo et al. (2012) as a horizontal
line, with the shaded area corresponding to the standard deviation.
}
\label{fig:ew}
\end{figure}

In order to quantitatively assess how emission lines change with 
inclination when blue-shifted absorption 
may affect the line profile, we define the `red wing equivalent width' ($W_{\lambda,RW}$) as
\begin{equation}
W_{\lambda,RW} = \int_{\lambda_0}^{\lambda^\prime} \left( 1 - \frac{F_\lambda}{F_0} \right) d\lambda
\label{rwew}
\end{equation}
where $F_0$ is the continuum flux and the integral is calculated from $\lambda_0$, line centre,
to a wavelength $\lambda^\prime$ where the flux has returned to the continuum level.
This quantity is shown as a function of inclination in figure 5 for the \civ\ and \mg\ UV lines.
We also show the $W_{\lambda,RW}$ expected from isotropic line emission and a foreshortened 
and limb darkened disc  as well as $1/2$ equivalent widths from \cite{dipompeo2012b}.

@@

It is also possible that the equatorial paradigm invoked from early polarisation
studies \citep{goodrich1995, cohen1995,brotherton2006} 
is an over-simplification, or is incorrect. 
High brightness temperatures in some RL BALQSOs imply polar outflows \citep{zhou2006} 
and \cite{bruni2012} find that
RL BALQSOs possess similar radio spectral indices to normal RL quasars, 
suggestive of comparable inclinations. 
In addition, \cite{marin2013} find a bending angle of $\sim45^\circ$ is required to 
explain the polarisation dichotomy of type 1 and 2 AGN using an Elvis-type wind 
model (Elvis 2000). It is therefore possible 
that type 1 quasars and BALQSOs are generally viewed from a fairly narrow range of less
equatorial angles ($\sim0$-$45^\circ$), or that {\em both} 
evolutionary and geometric explanations are required.
We suggest that future modelling should include predictions of polarisation signatures
from a detailed radiative transfer simulation, allowing direct 
comparison with spectropolarimetry of BALQSOs.
